---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

ANSWER HERE
The categories vary greatly in severity, from accident and robbery to homicide and assault. From a reporting standpoint, and assuming my audience is UMD students, some categories are more important to UMD students than others. Students would most likely be concerned about rates of assault, robbery, breaking and entering, and auto theft (for those who have cars). UMD students would also be more concerned about breaking and entering in residential buildings rather than commercial buildings. Crimes with multiple catagories can make summarizing the data more difficult because it doesn't allow you to generalize as much. 

### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
total_incidents <- pg_crime |> mutate(total_incidents=(Jan+Feb+Mar+Apr+May+Jun+Jul+Aug+Sep+Oct+Nov+Dec))

pg_crime_averages <- total_incidents |> mutate(pg_crime_averages=(total_incidents/12))

```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

Accident, auto stolen, and theft from auto are the three categories with the highest monthly average of incidents. Homicide, breaking and entering other, and sex offense are the three categories with the lowest average monthly average of incidents. There are many auto related incidents in PG County and not as many violent crimes. I am wondering why Robbery Residential has an NA value for March and October. I am assuming it is because there were zero reported incidents, but I am not sure. 


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(june_change = (`Jun`- `pg_crime_averages`)/`pg_crime_averages` * 100)
summer_crime_averages <- summer_crime_averages |> mutate(july_change = (`Jul`- `pg_crime_averages`)/`pg_crime_averages` * 100)
summer_crime_averages <- summer_crime_averages |> mutate(august_change = (`Aug`- `pg_crime_averages`)/`pg_crime_averages` * 100)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

Over the summer months, there are almost equal occurrences of a percentage increase and decrease in the average number of incidents per month of each crime category. Overall, the percentage increases are greater, which indicates a higher crime rate in the summer. Both Theft and Breaking and Entering Other had percentage decreases in each summer month. Auto Stolen, Homicide, and Theft From Auto had percentage increases in each summer month. 

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Auto thefts were up the summer of 2024 in Prince George's County, as all three summer months experienced higher rates than the monthly average. 

There were 4790 total reports of stolen vehicles in 2024, according to data provided by the Prince George's County Police Department. The monthly average number of incidents was 399. June expereinced a 10% increase, and July and August both had a 5% increase. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
