---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |> mutate (violent_rate_2019=violent2019/pop2019*2000, property_rate_2019=property2019/pop2019*2000, violent_rate_2020=violent2020/pop2020*2000, property_rate_2020=property2020/pop2020*2000 )
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

In 2019, Baltimore City, Ocean City, Williamsport, and Elkton had the four highest violent crime rates. Ocean City, Elkton, and Hyattsville had the highest property crime rates in 2019. In 2020, Ocean City had the highest violent crime rate and the highest property crime rate. In 2020, Baltimore City, Cambridge, and Princess Anne also had higher violent crime rates. For property crime in 2020, Colmar Manor, Landover Mills, and Oakland had higher rates. 

The story is that Ocean City's already high crime rate skyrocketed during 2020. It had a violent crime rate of 26 per 2000 that increased to 47 per 2000 in 2020. I also think it's interesting that Ocean City's crime rate was so high in general. Most people label Baltimore as a very dangerous city but don't treat Ocean City like it falls second-in-line and sometimes first. 


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate (violent_change=(violent_rate_2020-violent_rate_2019)/violent_rate_2019 * 100, property_change=(property_rate_2020-property_rate_2019)/property_rate_2019 * 100) 
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

Accident had one of each in 2019 and then two of each in 2020. Of the cities that had a percentage change result, most of them decreased in violent and property crimes. What I found more interesting is that some cities had a dramatic increase in violent and property crime rates from 2019-2020. The top three increases in violent crime rates were Edmonston (895%), Poolesville (607%) and Pocomoke City (301%). The top three increases in property crime rates were Crisfield (321%), Cecilton (200%) and Garett Park (188%). I find these outliers much interesting, especially the fact that they are all smaller towns  

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Violent crime rates decreased for many cities in Maryland from 2019 to 2020, but some of the highest increases occurred in smaller towns in the state. 

Edmonston experienced the highest violent crime rate increase of 895%, and Poolesville came in second with a 607% increase. Of the ten cities with the highest increase in violent crime rates, eight had a population in 2020 of under 5,000 people and classify as a small town. The two cities that had a population higher than 5,000 were Poolesville with 5191 residents and Glenarden with 6177 residents. Meanwhile, Baltimore and Frederick, the two most populous cities, experienced a decrease in violent rates of 13% and 95%, respectively. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
